### 2016-07-07 motivation.tex
Fixed Figs. 2 and 4.

\begin{figure}[!h]
\centering
\subcaptionbox {
  {\sc rc} Execution 1
  \label{fig:motiv-eg-1-a}
} [
  0.55\columnwidth
] {
  \includegraphics[scale=0.6]{Figures/motiv-eg-1-a}
}
%\hspace*{0.5in}
\subcaptionbox {
  {\sc rc} Execution 2
  \label{fig:motiv-eg-1-b}
}{
  \includegraphics[scale=0.6]{Figures/motiv-eg-1-b}
}
\caption{\small A possible execution of the program shown in Fig.~\ref{fig:motiv-eg-1} under
  \iso{Read Committed} isolation level.}
\label{fig:rc-ex}
\end{figure}

The figure depicts an execution as a series of read, write, and commit
operations. The subscript of an operation indicates the transaction
that executes it (Both transactions read \C{B} twice, hence two read
operations). In the execution on left, transaction \C{Wd1} reads the
current balance (\C{k}) and writes the new balance (\C{k-a1}), but
before it commits, transaction \C{Wd2} executes and commits, writing
the new balance (\C{k-a2}). {\sc rc} isolation prevent \C{Wd2} from
witnessing the uncommitted writes of transaction \C{Wd1}.
Subsequently committing \C{Wd1} leads to the loss of \C{Wd2}'s updates
(the so-called \emph{lost update anomaly}~\cite{berenson}), resulting
in an incorrect balance of \C{k-a1}. The execution on right describes
a similar scenario with \C{Wd1} and \C{Wd2} exchanging their roles.
Clearly, {\sc rc} is an excessively weak isolation level for this
program because it loses the updates of one transaction, resulting in
the violation of the post-condition.  A stronger isolation level that
prevents lost updates is required. \iso{Snapshot Isolation} ({\sc
si})~\cite{berenson} fits this requirement; {\sc si} effectively
serializes transactions that update a shared data object by aborting
and re-executing a transaction if write-write conflicts are detected
during its commit.\footnote{{\sc si} however does not serialize
transactions in the absence of write-write conflicts.} Since {\sc si},
unlike {\sc ser}, does not need expensive mechanisms such as
lock-based concurrency control or runtime monitoring, it is also more
efficient, making it appropriate for both withdraw transactions.

@@

\caption{\small $\hbZ$ arrows allowed by $\psi_{SI}$
 and $\psi_{RC}$ specifications
% (Figs.~\ref{fig:hb-eg-1},~\ref{fig:hb-eg-2},~\ref{fig:hb-eg-3},~\ref{fig:hb-eg-4})
for the example in Fig.~\ref{fig:motiv-eg-1}. Each
arrow represents an $\hbZ$ relationship between all operations of
\C{Wd2} and an operation in \C{Wd1}. Letters R, W and C stand for
read, write and commit, respectively.}
\label{fig:motiv-eg-1-hb}
\vspace*{-8pt}
\end{figure}

>>>>

\begin{figure}[!h]
\centering
\subcaptionbox {
  {\sc rc} Execution 1
  \label{fig:motiv-eg-1-a}
} [
  0.55\columnwidth
] {
  \includegraphics[scale=0.6]{Figures/motiv-eg-1-a}
}
%\hspace*{0.5in}
\subcaptionbox {
  {\sc rc} Execution 2
  \label{fig:motiv-eg-1-b}
}{
  \includegraphics[scale=0.6]{Figures/motiv-eg-1-b}
}
\caption{\small A possible execution of the program shown in Fig.~\ref{fig:motiv-eg-1} under
  \iso{Read Committed} isolation level. Transaction \C{Wd1} is shown
  against lighter green background, and transaction \C{Wd2} against
  darker red background. Each transaction reads the balance (\C{B})
  twice, hence two \C{Reads}.}
\label{fig:rc-ex}
\end{figure}

The figure depicts an execution as a series of read, write, and commit
operations. In the execution on left, transaction \C{Wd1} (green)
reads the current balance (\C{k}) and writes the new balance
(\C{k-a1}), but before it commits, transaction \C{Wd2} (red) executes
and commits, writing the new balance (\C{k-a2}). {\sc rc} isolation
prevent \C{Wd2} from witnessing the uncommitted writes of transaction
\C{Wd1}.  Subsequently committing \C{Wd1} leads to the loss of
\C{Wd2}'s updates (the so-called \emph{lost update
anomaly}~\cite{berenson}), resulting in an incorrect balance of
\C{k-a1}. The execution on right describes a similar scenario with
\C{Wd1} and \C{Wd2} exchanging their roles.  Clearly, {\sc rc} is an
excessively weak isolation level for this program because it loses the
updates of one transaction, resulting in the violation of the
post-condition.  A stronger isolation level that prevents lost updates
is required. \iso{Snapshot Isolation} ({\sc si})~\cite{berenson} fits
this requirement; {\sc si} effectively serializes transactions that
update a shared data object by aborting and re-executing a transaction
if write-write conflicts are detected during its commit.\footnote{{\sc
si} however does not serialize transactions in the absence of
write-write conflicts.} Since {\sc si}, unlike {\sc ser}, does not
need expensive mechanisms such as lock-based concurrency control or
runtime monitoring, it is also more efficient, making it appropriate
for both withdraw transactions.

@@

\caption{\small For the example in Fig.~\ref{fig:motiv-eg-1},
$\psi_{SI}$ specification allows $\hbZ$ arrows (hence the execution)
in Fig.~\ref{fig:hb-eg-1}. $\psi_{RC}$ specification allows
Figs.~\ref{fig:hb-eg-1},~\ref{fig:hb-eg-2},~\ref{fig:hb-eg-3}, and
~\ref{fig:hb-eg-4}. Each arrow represents an $\hbZ$ relationship
between all operations of \C{Wd2} and an operation in \C{Wd1}. Letters
R, W and C stand for read, write and commit, respectively.}
\label{fig:motiv-eg-1-hb}
\vspace*{-8pt}
\end{figure}
